/***********************************************
** DATA ENGINEERING PLATFORMS (MSCA 31012)
** File: RMySQL
** Desc: Connecting R to MySQL database
** Auth: Shreenidhi Bharadwaj
** Date: 02/06/2019
** ALL RIGHTS RESERVED | DO NOT DISTRIBUTE
************************************************/

We can connect to MySQL using the RMySQL package.
```{r}
install.packages("RMySQL");
install.packages("ggplot2");
```

Initialize libraries
```{r, echo=FALSE}
library(RMySQL);
library(ggplot2);
library(DBI);
```

Set up the connection
```{r, echo=FALSE}
connection = dbConnect(MySQL(),user="root", password="rootroot", 
                       dbname="classicmodels", host="localhost");
```

Now get some data
```{r}
myQuery <- "SELECT COUNT(*) FROM offices;"
dbGetQuery(connection, myQuery)
```


Find all the orders that are inprocess or have been shipped 
```{r}
myQuery <- 
"SELECT 
    orderNumber, orderDate, requiredDate, customerNumber, status
FROM
    orders
WHERE
    (status = 'In Process'
        OR status = 'shipped');"
dbGetQuery(connection, myQuery);
```

# arithmetic mean of the buyPrice for all products
```{r}
myQuery <- 
"SELECT 
    VARIANCE(quantityOrdered) AS varQuantityOrdered
FROM
    orderdetails;"
dbGetQuery(connection, myQuery);
```

# number of orders and status of orders
```{r}
myQuery <- 
"SELECT 
    customerNumber, 
    status, 
    COUNT(*) AS numberOfOrders
FROM
    orders
GROUP BY customerNumber , status
ORDER BY numberOfOrders DESC;"

myData <- dbGetQuery(connection, myQuery);
ggplot(myData, aes(status, numberOfOrders)) + geom_bar(stat="identity")

```


Close active DB Connections

```{r}
dbGetQuery(connection, "show processlist")

dbDisconnectAll <- function()
{
  ile <- length(dbListConnections(MySQL())  )
  lapply( dbListConnections(MySQL()), function(x) dbDisconnect(x) )
  cat(sprintf("%s connection(s) closed.\n", ile))
}

dbDisconnectAll()

```

```{r}
dbconnect = dbConnect(MySQL(),user="root", password="rootroot", 
                dbname="classicmodels", host="localhost")
```

```{sql, connection=dbconnect}
SELECT count(*) FROM customers;
```
```{sql, connection=dbconnect}
# List all products that matches a pattern ( string )
SELECT
    productCode, 
    SUM(quantityOrdered) as totalQuantity, 
    SUM(quantityOrdered*priceEach) as totalPrice
FROM
    orderdetails
WHERE productCode like 'S18%'
GROUP BY productCode 
/* ASC */
HAVING totalPrice > 10000
ORDER BY totalPrice DESC
LIMIT 5;
```

# Additional References
# https://programminghistorian.org/en/lessons/getting-started-with-mysql-using-r
